---
title: "B563_hw3"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1

```{r}
x = runif(n=1000,min=-4,max=4)
y = -x
z = rep.int(1,1000)
X = data.frame(x,y,z)

cov = cov(X)
svd = svd(cov)
U = svd$u
D = diag(svd$d)
U
D
```

## 2(a)

```{r}
library("mosaic")

df =  read.csv("D:/Data Mining/hw3/Arrests.csv")
head(df)

t = table(df[,2],df[,3])
t
p_br = t[2,1]/sum(t[2,1]+t[2,2])
p_br
p_wr = t[2,2]/sum(t[2,1]+t[2,2])
p_wr

mosaicplot(table(df[,3],df[,2]),xlab="colour",ylab="released")


```

## 2(b)

- It is evident that given employment status, colour and released are not independent. Heights of blacks and white given employment status and released status is not same.

```{r}

df1 = df[df$employed == 'Yes',]
df2 = df[df$employed == 'No',]
head(df1)


t1 = table(df1[,3],df1[,2])
t1
t2 = table(df2[,3],df2[,2])
t2

tab1 = mosaicplot(t1)
tab2 = mosaicplot(t2)

```



## 2(c)

- It is evident that given checks, colour and released are not independent. Heights of blacks and white given check status and released status is not same.

```{r}
df_1 = df[df$checks == 0,]
df_2 = df[df$checks == 1,]
df_3 = df[df$checks == 2,]
df_4 = df[df$checks == 3,]
df_5 = df[df$checks == 4,]
df_6 = df[df$checks == 5,]
df_7 = df[df$checks == 6,]

t1 = table(df1[,3],df1[,2])
t1
t2 = table(df2[,3],df2[,2])
t2

tab1 = mosaicplot(table(df_1[,3],df_1[,2]))
tab2 = mosaicplot(table(df_2[,3],df_2[,2]))
tab3 = mosaicplot(table(df_3[,3],df_3[,2]))
tab4 = mosaicplot(table(df_4[,3],df_4[,2]))
tab5 = mosaicplot(table(df_5[,3],df_5[,2]))
tab6 = mosaicplot(table(df_6[,3],df_6[,2]))
tab7 = mosaicplot(table(df_7[,3],df_7[,2]))

```


# 2(d) 
- Given employment, we can see that whites are more probable to get released than blacks
- For checks also, we can see that whites are more probable to get released than blacks except when check = 7
- It is evident that higher arrest rate is there among blacks and data appears to be consistent with racial bias
- Also, there could be external factors which may affect the higher arrest rate of blacks but it cannot be explained through mosaic plot 


# 3(a)

```{r}
a = rexp(1000,1)
plot(ecdf(a))


```

# 3(b)
- The distribution is uniform

```{r}
b = ecdf(a)
b1 = b(sort(a))
plot(b1)

hist(b1)


```

# 3(c)

- Assumed two distribution, one of which is exponentially distributed and other random distributed with each has 10 datapoints
- F(X<x(i)) = (number of X less than x(i))/1000 where i is the index
- For any given index i , F(X<x(i)) will have same probablity irrespective of value of i 
- Thus, P(x) will have same distribution irrespective of any values of i

```{r}

a1 = runif(1000)
b2 = ecdf(a1)
b3 = b2(sort(a1))
plot(b3)

d = density(b3)
plot(d)

d1 = density(b1)
plot(d1)
```


## 4(a)
- The different models of time series has different amplitude and frequency. 
- Model 1: It has low frequency and high amplitude
- Model 2: It has high frequency and low amplitude
- Model 3: It has high frequency and high amplitude
- Model 4 : It has low frequency and low  amplitude

```{r}

library(reshape2)
library(ggplot2)
library(forecast)
library(fpp2)
library(ggplot2)
library(tidyr)

df =  read.csv("D:/Data Mining/hw3/time_series.csv")
head(df)

df_t = t(df)
cov = cov(df_t)
svd = svd(cov)
U = svd$u

X = df_t%*%U[,1:4]

ts.plot(X,gpars=list(col=rainbow(10)))



```


# 4(b)

```{r}
library("forecast")
amp = apply(df, 1, max)

df_t = t(df)

freq = rep(NA,200)
for (i in 1:ncol(df_t))
  freq[i] = findfrequency(df_t[,i])

plt = plot(freq,amp)
plt

```




